---
title: "Compare versions (Water Availability)"
author: "Liam D. Bailey"
format: html
editor: source
---

```{r}
#| echo: false
#| include: false

## SETUP
if (!require("pacman")) install.packages("pacman")
pacman::p_load(here, dplyr, magrittr, sf, terra, stars, exactextractr, skimr, readr, tidyr, openxlsx, mapview, ggplot2)
sf::sf_use_s2(FALSE) # due to issues in HydroBASINS geometries

country_borders <- read_sf(
  file.path(
    "C:\\Users\\liam.bailey\\WWF-Deutschland\\FB-WuM-Daten - Water & Biodiversity Stewardship",
    "10_RiskSuite", "RFS_New_Folders", "05_Data_and_Analysis",
    "01_Data", "Borders_WorldBank", "World Bank Official Boundaries - Admin 0.gpkg"
  )
) |> 
  sf::st_simplify(dTolerance = 0.01)
```

## Introduction

:::{.callout-note}
- Unified layer v1.1 updates to use Aqueduct 4.0 and still usese Blue Water Scarcity (Mekonnen & Hoekstra).
- Unified layer v2.0 replaces Blue Water Scarcity (Mekonnen & Hoekstra) with [Freshwater Quantity Targets v1.0 (Hogeboom et al. 2024)](https://doi.org/10.4121/089afa18-bbd5-493e-8a6c-39e1ca7b2a59.v1).
- Unified layer v2.1 uses newly released [Freshwater Quantity Targets v2.0 (Hogeboom et al. 2024)](https://doi.org/10.4121/089afa18-bbd5-493e-8a6c-39e1ca7b2a59.v2).
- We test the following questions:
1. How do unified layers v1.1, v2.0, and v2.1 differ from one another?
2. How do risks differ between Blue Water Scarcity, Freshwater Quantity Targets v1.0, and Freshwater Quantity Targets v2.0?
3. How do conclusions differ when using Freshwater Quantity Targets v2.0 v. unified layer?
4. Which water availability datasets are the drivers of risk (i.e. which dataset has the highest risk in each location)?
:::

## Load data

Create a single (simplified) layer with data from all three versions

```{r}
unified2.1 <- read_sf(here("v2.1", "outputs", "sbtn_son_water_v2.1.shp")) |> 
  st_simplify(dTolerance = 0.01) 

unified2.0 <- read_sf(here("v2.0", "outputs", "sbtn_son_water_v2.0.shp")) |>
  sf::st_drop_geometry()

unified1.1 <- read_sf(here("v1.1", "outputs", "sbtn_son_water_v1.1.shp")) |>
  sf::st_drop_geometry()

allunified <- unified2.1  |> 
  select(HYBAS_ID, fqt_raw, fqt_n, wa_max_2.1 = wa_max) |> 
  left_join(unified2.0 |> select(HYBAS_ID, wa_max_2.0 = wa_max, fqt_n_v1.0 = fqt_n), by = "HYBAS_ID") |> 
  left_join(unified1.1 |> select(HYBAS_ID, wa_max_1.1 = wa_max, wsb_n), by = "HYBAS_ID")
```

## How do risks differ between Blue Water Scarcity, Freshwater Quantity Targets v1.0, and Freshwater Quantity Targets v2.0?

```{r}
BlueWaterScarcity <- ggplot() +
  geom_sf(data = allunified,
          aes(fill = as.factor(wsb_n)), colour = NA) +
  geom_sf(data = country_borders,
          fill = NA, colour = "black", linewidth = 0.1) +
  labs(title = "Blue Water Scarcity") +
  scale_fill_manual(values = c("#FFFF99", "#FFE600", "#FF9900", "#FF1900", "#990000"),
                    name = "Blue Water Scarcity risk") +
  theme_void() +
  theme(legend.position = "bottom",
        legend.title.position = "top",
        legend.key.width = unit(0.5, "in"))

FQTv1.0 <- ggplot() +
  geom_sf(data = allunified,
          aes(fill = as.factor(fqt_n_v1.0)), colour = NA) +
  geom_sf(data = country_borders,
          fill = NA, colour = "black", linewidth = 0.1) +
  labs(title = "Freshwater Quantity Target (v1.0)") +
  scale_fill_manual(values = c("#FFFF99", "#FFE600", "#FF9900", "#FF1900", "#990000"),
                    name = "Freshwater Quantity Target risk") +
  theme_void() +
  theme(legend.position = "bottom",
        legend.title.position = "top",
        legend.key.width = unit(0.5, "in"))

FQTv2.0 <- ggplot() +
  geom_sf(data = allunified,
          aes(fill = as.factor(fqt_n)), colour = NA) +
  geom_sf(data = country_borders,
          fill = NA, colour = "black", linewidth = 0.1) +
  labs(title = "Freshwater Quantity Target (v2.0)") +
  scale_fill_manual(values = c("#FFFF99", "#FFE600", "#FF9900", "#FF1900", "#990000"),
                    name = "Freshwater Quantity Target risk") +
  theme_void() +
  theme(legend.position = "bottom",
        legend.title.position = "top",
        legend.key.width = unit(0.5, "in"))

combo <- BlueWaterScarcity + FQTv1.0 + FQTv2.0 + patchwork::plot_layout(nrow = 1, ncol = 3)

ggsave(plot = combo, filename = here("outputs", "plots", "hogeboom_check", "single_data_compare.png"), width = 15, height = 5)
```

## How do unified layers v1.1, v2.0, and v2.1 differ from one another?

```{r}
plot_data <- allunified |> 
  mutate(diff_2.0 = factor(wa_max_2.0 - wa_max_1.1),
         diff_2.1 = factor(wa_max_2.1 - wa_max_1.1))

v1.1 <- ggplot() +
  geom_sf(data = plot_data,
          aes(fill = as.factor(wa_max_1.1)), colour = NA) +
  geom_sf(data = country_borders,
          fill = NA, colour = "black", linewidth = 0.1) +
  labs(title = "Water Availability risk (Unified layer v1.1)") +
  scale_fill_manual(values = c("#FFFF99", "#FFE600", "#FF9900", "#FF1900", "#990000"), name = "Water Availability riks (v1.1)") +
  theme_void() +
  theme(legend.position = "bottom",
        legend.title.position = "top",
        legend.key.width = unit(0.5, "in"))

v2.0_diff <- ggplot() +
  geom_sf(data = plot_data,
          aes(fill = diff_2.0), colour = NA) +
  geom_sf(data = country_borders,
          fill = NA, colour = "black", linewidth = 0.1) +
  labs(title = "Water Availability risk differences (v1.1 and v2.0)") +
  scale_fill_manual(values = colorRampPalette(c(scales::muted("blue"), "white", scales::muted("red")), space = "Lab")(9), name = "Risk difference (v1.1 and v2.0)") +
  theme_void() +
  theme(legend.position = "bottom",
        legend.title.position = "top",
        legend.key.width = unit(0.5, "in"))

v2.1_diff <- ggplot() +
  geom_sf(data = plot_data,
          aes(fill = diff_2.1), colour = NA) +
  geom_sf(data = country_borders,
          fill = NA, colour = "black", linewidth = 0.1) +
  labs(title = "Water Availability risk differences (v1.1 and v2.1)") +
  scale_fill_manual(values = colorRampPalette(c(scales::muted("blue"), "white", scales::muted("red")), space = "Lab")(9), name = "Risk difference (v1.1 and v2.1)") +
  theme_void() +
  theme(legend.position = "bottom",
        legend.title.position = "top",
        legend.key.width = unit(0.5, "in"))

combo <- v1.1 + v2.0_diff + v2.1_diff + patchwork::plot_layout(nrow = 1, ncol = 3)

ggsave(plot = combo, filename = here("outputs", "plots", "hogeboom_check", "unified_layer_compare.png"), width = 15, height = 5)
```

## How do conclusions differ when using Freshwater Quantity Targets v2.0 v. unified layer?

```{r}
plot_data <- allunified |> 
  mutate(diff_v2.0 = as.factor(fqt_n_v1.0 - wa_max_2.0),
         diff_v2.1 = as.factor(fqt_n - wa_max_2.1))

onlyFQT_v2.0 <- ggplot() +
  geom_sf(data = plot_data,
          aes(fill = diff_2.0), colour = NA) +
  geom_sf(data = country_borders,
          fill = NA, colour = "black", linewidth = 0.1) +
  labs(title = "Risk differences (FQT v1.0 v. unified layer v2.0)") +
  scale_fill_manual(values = colorRampPalette(c(scales::muted("blue"), "white", scales::muted("red")), space = "Lab")(9), name = "Risk difference") +
  theme_void() +
  theme(legend.position = "bottom",
        legend.title.position = "top",
        legend.key.width = unit(0.5, "in"))

onlyFQT_v2.1 <- ggplot() +
  geom_sf(data = plot_data,
          aes(fill = diff_2.1), colour = NA) +
  geom_sf(data = country_borders,
          fill = NA, colour = "black", linewidth = 0.1) +
  labs(title = "Risk differences (FQT v2.0 v. unified layer v2.1)") +
  scale_fill_manual(values = colorRampPalette(c(scales::muted("blue"), "white", scales::muted("red")), space = "Lab")(9), name = "Risk difference") +
  theme_void() +
  theme(legend.position = "bottom",
        legend.title.position = "top",
        legend.key.width = unit(0.5, "in"))

combo <- onlyFQT_v2.0 + onlyFQT_v2.1 + patchwork::plot_layout(nrow = 1, ncol = 2)

ggsave(plot = combo, filename = here("outputs", "plots", "hogeboom_check", "FQT_alone.png"), width = 10, height = 5)
```

## Which water availability datasets are the drivers of risk (i.e. which dataset has the highest risk in each location)?

```{r}
combo_data <- son_water_keydata_wa |> 
  group_by(HYBAS_ID) |> 
  summarise(top_mod = case_when(combo ~ "Multiple",
                                BWS_top ~ "BWS",
                                WDP_top ~ "WDP",
                                FQT_top ~ "FQT")) |> 
  ungroup()

combo_data_sf <- son_water_keydata_wa_sf |> 
  left_join(combo_data, by = "HYBAS_ID")
```

```{r}
top_model_plot <- ggplot() +
  geom_sf(data = combo_data_sf,
          aes(fill = top_mod), colour = NA) +
  scale_fill_manual(values = c("#66c2a5", "#fc8d62", "grey90", "#8da0cb")) +
  theme_void() +
  theme(panel.background = element_rect(fill = "white", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA),
        legend.position = "bottom")

ggsave(plot = top_model_plot, filename = here("./v2.1/outputs", "all_drivers.png"))
```