---
title: "SBTN State of Nature Water unified layers"
subtitle: "water availability and water pollution - National and Sub-national level, weighted by sector"
author: "Rafael Camargo"
date: "Jun 25, 2024"
engine: knitr
format:
  html:
    toc: true
execute:
  warning: false    
---

## 1. Setup

Load required packages

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(here, dplyr, readr, readxl, sf, tidyr, terra, fasterize, purrr, exactextractr, skimr, mapview, openxlsx)
```

```{r}
sharepoint_dir <- "C:\\Users\\Liam.Bailey\\wwfgermany"
temp_raster_dir <- here::here("/outputs", "temp_rasters")
dir.create(temp_raster_dir)
```

## 2. Load data

```{r}
## We want to read in data at h6 level
## No need to use h7 level for this aggregation
son_water <- read_rds(here("/outputs", "h6", "sbtn_son_water_v2.rds"))
```

```{r}
aqueduct_adm0 <- read_excel(file.path(sharepoint_dir, "FB-OuR-Daten - WRF", "ws_gis", "1_library", "aqueduct_4.0", "aqueduct-4-0-country-rankings", "Aqueduct40_rankings_download_Y2023M07D05", "Aqueduct40_rankings_download_Y2023M07D05.xlsx"),
                            sheet = "country_baseline")
```

```{r}
aqueduct_adm1 <- read_excel(file.path(sharepoint_dir, "FB-OuR-Daten - WRF", "ws_gis", "1_library", "aqueduct_4.0", "aqueduct-4-0-country-rankings", "Aqueduct40_rankings_download_Y2023M07D05", "Aqueduct40_rankings_download_Y2023M07D05.xlsx"),
                            sheet = "province_baseline")
```

Also read in GADM (v3.6) data and join with Aqueduct data. This helps deal with corruption/encoding issues in Aqueduct province names.

```{r}
GADM_adm1_sf <- read_sf(file.path(sharepoint_dir, "FB-OuR-Daten - WRF", "ws_gis", "1_library", "boundaries_gadm", "gadm36_Prov.shp"))

aqueduct_adm1 <- aqueduct_adm1 |> 
  rename(NAME_1_WRI = name_1) |> 
  left_join(GADM_adm1_sf |> 
              st_drop_geometry() |> 
              select(GID_0, GID_1, NAME_1),
            by = c("gid_0" = "GID_0",
                   "gid_1" = "GID_1"))
```

*National level administrative division*

```{r}
## We need to use OUR ESRI updated boundaries for RFS
## Rather than GADM borders that Rafael used
adm0 <- st_read(file.path(sharepoint_dir, "FB-OuR-Daten - WRF", "ws_gis", "1_library", "boundaries_esri", "world_adm0_esri_edited_diss.shp")) |> 
  ## Renamed to match column in WRI data
  rename(GID_0 = ISO_3DIGIT)
```

*Sub-national level administrative division*

```{r}
## We need to use OUR ESRI updated boundaries for RFS
## Rather than GADM borders that Rafael used
adm1 <- st_read(file.path(sharepoint_dir, "FB-OuR-Daten - WRF", "ws_gis", "1_library", "boundaries_esri", "world_adm1_esri_edited_diss.shp"))
```

## 3. Rasterize SoN water layers

```{r}
## Function to fasterize input data
fasterize_son_water <- function(var){
  fasterize(son_water, raster(son_water, res = 0.01), field = var, fun = "max") |>
    writeRaster(file.path(temp_raster_dir, paste0(var, ".tif")), overwrite = TRUE)
}

## Check if rasters are already created
current_rasters <- gsub(x = list.files(temp_raster_dir, pattern = ".tif"), replacement = "", pattern = ".tif")
missing_rasters <- c("wdp_n", "wsb_n", "cep_n", "nox_n", "pgp_n")[!c("wdp_n", "wsb_n", "cep_n", "nox_n", "pgp_n") %in% current_rasters]

## Create rasters for any files that are missing
if (length(missing_rasters) > 0){
  map(list(missing_rasters), fasterize_son_water) 
}
```

```{r}
list_wa <- list(
  rast(file.path(temp_raster_dir, "wdp_n.tif")),
  rast(file.path(temp_raster_dir, "wsb_n.tif"))
) |> 
  set_names("wdp_n", "wsb_n")
```

```{r}
list_wp <- list(
  rast(file.path(temp_raster_dir, "cep_n.tif")),
  rast(file.path(temp_raster_dir, "nox_n.tif")),
  rast(file.path(temp_raster_dir, "pgp_n.tif"))
) |> 
  set_names("cep_n", "nox_n", "pgp_n")
```

## 4. Prepare data to use as weights

### 4.1. WaterGap3 water consumption as weights for water depletion (provided by the authors)

```{r}
waterGap3_consumption <- read_csv(
  file.path(sharepoint_dir, "FB-OuR-Daten - 10_RiskSuite", "Data_and_Analysis", "01_Data", "WaterDepletion_WaterGap3_2000", "SectorConsumption_m3_annual.csv"),
  col_names = c("subBas_ID", "domestic", "manufacturing", "electric", "irrigation", "livestock")
)

waterGap3_consumption <- read_sf(file.path(sharepoint_dir,
                                           "FB-OuR-Daten - 10_RiskSuite", "Data_and_Analysis",
                                           "01_Data", "WaterDepletion_WaterGap3_2000", "WaterGap3Shape",
                                           "WaterGap3_AllBasins.shp")) |>
  dplyr::select(subBas_ID = ContUID) |>
  inner_join(waterGap3_consumption, by = "subBas_ID") |>
  mutate(
    industrial = manufacturing + electric,
    total = domestic + irrigation + livestock + industrial
  ) |>
  dplyr::select(
    dom = domestic,
    ind = industrial,
    irr = irrigation,
    liv = livestock,
    tot = total
  )

```

```{r}
## Function to fasterize data
fasterize_vars <- function(var){
  fasterize(waterGap3_consumption, raster(waterGap3_consumption, res = 0.01), field = var, fun = "max") |>
    writeRaster(file.path(temp_raster_dir, paste0("wdp_weight_", var, ".tif")), overwrite = TRUE)
}

## Check if rasters are already created
current_rasters <- gsub(x = list.files(temp_raster_dir, pattern = ".tif"), replacement = "", pattern = "wdp_weight_|.tif")
missing_rasters <- c("dom", "ind", "irr", "liv", "tot")[!c("dom", "ind", "irr", "liv", "tot") %in% current_rasters]

## Create rasters for any files that are missing
if (length(missing_rasters) > 0){
  map(list(missing_rasters), fasterize_vars) 
}
```

```{r}
wdp_weights <- list.files(temp_raster_dir, pattern = "wdp_weight_",
                          recursive = TRUE, full.names = TRUE) |> 
  map(rast) |> 
  set_names("dom", "ind", "irr", "liv", "tot")
```

### 4.2. Water Footprint Network Blue water footprint as weights for blue water scarcity (provided by the authors)

```{r}
wsb_weights <- list(
  rast(file.path(sharepoint_dir, "FB-OuR-Daten - 10_RiskSuite", "Data_and_Analysis", "01_Data", "BlueWaterFootprint_MekonnenHoekstra_2016", "WFbl_AgDomInd", "WFbl_m3yr", "wfbl_agr_m3y", "w001001.adf")),
  rast(file.path(sharepoint_dir, "FB-OuR-Daten - 10_RiskSuite", "Data_and_Analysis", "01_Data", "BlueWaterFootprint_MekonnenHoekstra_2016", "WFbl_AgDomInd", "WFbl_m3yr", "wfbl_dom_m3y", "w001001.adf")),
  rast(file.path(sharepoint_dir, "FB-OuR-Daten - 10_RiskSuite", "Data_and_Analysis", "01_Data", "BlueWaterFootprint_MekonnenHoekstra_2016", "WFbl_AgDomInd", "WFbl_m3yr", "wfbl_ind_m3y", "w001001.adf"))
) |> 
  set_names("agr", "dom", "ind")

## Add in the total footprint of all industries
wsb_weights$tot <- terra::rast(
  ## Create a stacked raster...
  list(
    wsb_weights$agr,
    ## Ensure that extent and resolution is same for all layers
    ## Need to coerce to same extent because agr has different ext
    terra::resample(wsb_weights$dom,
                    wsb_weights$agr),
    terra::resample(wsb_weights$ind,
                    wsb_weights$agr)
  )) |> 
  ## Sum all layers together
  terra::app(sum)
```

### 4.3. Global Cropland Extent (2019) [Potapov et al. 2021](https://glad.umd.edu/dataset/croplands)

```{r}
wp_weights <- rast(here(sharepoint_dir, "FB-OuR-Daten - 10_RiskSuite", "Data_and_Analysis", "01_Data", "CroplandExtent_Potapov_2021", "Global_cropland_3km_2019.tif"))
```

### 4.4. N and P fertilizer use as weights for water pollution indicators [Lu & Tian 2016](https://doi.pangaea.de/10.1594/PANGAEA.863323)

```{r}
n_use <- rast(file.path(sharepoint_dir, "FB-OuR-Daten - 10_RiskSuite", "Data_and_Analysis", "01_Data", "NPFertilizer_LuTian_2016", "Nfer_ASCII", "nfery2013.asc"))
```

```{r}
p_use <- rast(file.path(sharepoint_dir, "FB-OuR-Daten - 10_RiskSuite", "Data_and_Analysis", "01_Data", "NPFertilizer_LuTian_2016", "Pfer_ASCII", "pfery2013.asc"))
```

*Recode outliers*

```{r}
n_cap <- quantile(values(n_use), probs = 0.9998, na.rm = TRUE)
p_cap <- quantile(values(p_use), probs = 0.9998, na.rm = TRUE)
```

```{r}
n_use[n_use > n_cap] <- n_cap
p_use[p_use > p_cap] <- p_cap
```

*Normalise datasets and put N & P together*

```{r}
min_max_scale <- function(raster) {
  min_val <- min(values(raster), na.rm = TRUE)
  max_val <- max(values(raster), na.rm = TRUE)
  (raster - min_val) / (max_val - min_val)
}
```

```{r}
np_weights <- app(c(min_max_scale(n_use), min_max_scale(p_use)), fun = mean, na.rm = TRUE)
```

## 5. Spatial aggregation to administrative divisions

```{r}
aggregate_to_adm <- function(indicator, weight_by, key, to_adm, adm_id){
  exact_extract(indicator, to_adm, fun = "weighted_mean", weights = resample(weight_by, indicator), default_weight = 0, append_cols = adm_id, progress = TRUE) |> 
    rename({{key}} := weighted_mean)
} 
```

### 5.1. National level

```{r}
bws_adm0 <- aqueduct_adm0 |> 
  filter(indicator_name == "bws") |> 
  mutate(
    weight = if_else(weight == "One", "n", tolower(weight)),
    score = na_if(score, -9999),
    # rescale values 0-5 to 1-5 applying (new_max - new_min) * (x - min_old) / (max_old - min_old) + min_new
    score = (5-1) * ((score-0) / (5-0)) + 1 
  ) |> 
  select(GID_0 = gid_0, weight, score) |> 
  pivot_wider(names_from = weight, values_from = score, names_prefix = "bws_") |> 
  relocate(GID_0, bws_n, bws_dom, bws_ind, bws_irr, bws_liv, bws_tot)

## Don't try and fill in disputed areas...this is too complex
# bws_adm0 <- bws_adm0 |> 
#   bind_rows(
#     filter(bws_adm0, GID_0 == "IND") |> mutate(GID_0 = "Z01"),
#     filter(bws_adm0, GID_0 == "CHN") |> mutate(GID_0 = "Z02"),
#     filter(bws_adm0, GID_0 == "CHN") |> mutate(GID_0 = "Z03"),
#     filter(bws_adm0, GID_0 == "IND") |> mutate(GID_0 = "Z04"),
#     filter(bws_adm0, GID_0 == "IND") |> mutate(GID_0 = "Z05"),
#     filter(bws_adm0, GID_0 == "PAK") |> mutate(GID_0 = "Z06"),
#     filter(bws_adm0, GID_0 == "IND") |> mutate(GID_0 = "Z07"),
#     filter(bws_adm0, GID_0 == "CHN") |> mutate(GID_0 = "Z08"),
#     filter(bws_adm0, GID_0 == "IND") |> mutate(GID_0 = "Z09")
#   )
```

```{r}
son_water_adm0 <- adm0 |>
  # baseline water stress (bws)
  left_join(bws_adm0, by = "GID_0") |> 
  # water depletion (wdp)
  left_join(exact_extract(list_wa$wdp_n, adm0, fun = "mean", append_cols = "GID_0", progress = TRUE) |> rename(wdp_n = mean), by = "GID_0") |>
  left_join(aggregate_to_adm(list_wa$wdp_n, wdp_weights$dom, wdp_dom, adm0, "GID_0"), by = "GID_0") |>
  left_join(aggregate_to_adm(list_wa$wdp_n, wdp_weights$ind, wdp_ind, adm0, "GID_0"), by = "GID_0") |>
  left_join(aggregate_to_adm(list_wa$wdp_n, wdp_weights$irr, wdp_irr, adm0, "GID_0"), by = "GID_0") |>
  left_join(aggregate_to_adm(list_wa$wdp_n, wdp_weights$liv, wdp_liv, adm0, "GID_0"), by = "GID_0") |>
  left_join(aggregate_to_adm(list_wa$wdp_n, wdp_weights$tot, wdp_tot, adm0, "GID_0"), by = "GID_0") |>
  # blue water scarcity (wsb)
  left_join(exact_extract(list_wa$wsb_n, adm0, fun = "mean", append_cols = "GID_0", progress = TRUE) |> rename(wsb_n = mean), by = "GID_0") |>
  left_join(aggregate_to_adm(list_wa$wsb_n, wsb_weights$agr, wsb_agr, adm0, "GID_0"), by = "GID_0") |>
  left_join(aggregate_to_adm(list_wa$wsb_n, wsb_weights$dom, wsb_dom, adm0, "GID_0"), by = "GID_0") |>
  left_join(aggregate_to_adm(list_wa$wsb_n, wsb_weights$ind, wsb_ind, adm0, "GID_0"), by = "GID_0") |>
  left_join(aggregate_to_adm(list_wa$wsb_n, wsb_weights$tot, wsb_tot, adm0, "GID_0"), by = "GID_0") |> 
  # coastal eutrophication potential (cep)
  left_join(exact_extract(list_wp$cep_n, adm0, fun = "mean", append_cols = "GID_0", progress = TRUE) |> rename(cep_n = mean), by = "GID_0") |>
  left_join(aggregate_to_adm(list_wp$cep_n, wp_weights, cep_agr, adm0, "GID_0"), by = "GID_0")|>
  left_join(aggregate_to_adm(list_wp$cep_n, np_weights, cep_np, adm0, "GID_0"), by = "GID_0") |>
  # nitrate-nitrite concentration (nox)
  left_join(exact_extract(list_wp$nox_n, adm0, fun = "mean", append_cols = "GID_0", progress = TRUE) |> rename(nox_n = mean), by = "GID_0") |>
  left_join(aggregate_to_adm(list_wp$nox_n, wp_weights, nox_agr, adm0, "GID_0"), by = "GID_0") |>
  left_join(aggregate_to_adm(list_wp$nox_n, np_weights, nox_np, adm0, "GID_0"), by = "GID_0") |>
  # periphyton growth potential (pgp)
  left_join(exact_extract(list_wp$pgp_n, adm0, fun = "mean", append_cols = "GID_0", progress = TRUE) |> rename(pgp_n = mean), by = "GID_0") |>
  left_join(aggregate_to_adm(list_wp$pgp_n, wp_weights, pgp_agr, adm0, "GID_0"), by = "GID_0") |>
  left_join(aggregate_to_adm(list_wp$pgp_n, np_weights, pgp_np, adm0, "GID_0"), by = "GID_0")|>
  
  mutate(across(where(is.numeric), ~ round(.x, 1))) |> 
  rowwise() |> 
  mutate(
    wa_max_n = max(c(bws_n, wdp_n, wsb_n), na.rm = TRUE),
    wa_max_dom = max(c(bws_dom, wdp_dom, wsb_dom), na.rm = TRUE),
    wa_max_ind = max(c(bws_ind, wdp_ind, wsb_ind), na.rm = TRUE),
    wa_max_irr = max(c(bws_irr, wdp_irr, wsb_agr), na.rm = TRUE),
    wa_max_liv = max(c(bws_liv, wdp_liv, wsb_n), na.rm = TRUE),
    wa_max_tot = max(c(bws_tot, wdp_tot, wsb_tot), na.rm = TRUE),
    wp_max_n = max(c(cep_n, nox_n, pgp_n), na.rm = TRUE),
    wp_max_agr = max(c(cep_agr, nox_agr, pgp_agr), na.rm = TRUE),
    wp_max_np = max(c(cep_np, nox_np, pgp_np), na.rm = TRUE)
  ) |> 
  ungroup() |> 
  relocate(wa_max_n:wa_max_tot, .before = bws_n) |>
  relocate(wp_max_n:wp_max_agr, .before = cep_n) |> 
  relocate(geometry, .after = last_col()) |> 
  mutate(across(where(is.numeric), ~ if_else(between(.x, 1, 5), .x, NA)))
```

```{r}
st_drop_geometry(son_water_adm0) |> 
  skim()
```

### 5.2. Sub-National level

```{r}
bws_adm1 <- aqueduct_adm1 |> 
  filter(indicator_name == "bws") |> 
  mutate(
    weight = if_else(weight == "One", "n", tolower(weight)),
    score = na_if(score, -9999),
    # rescale values 0-5 to 1-5 applying (new_max - new_min) * (x - min_old) / (max_old - min_old) + min_new
    score = (5-1) * ((score-0) / (5-0)) + 1 
  ) |> 
  select(GID_0 = gid_0, GID_1 = gid_1, NAME_0 = name_0, NAME_1, weight, score) |> 
  pivot_wider(names_from = weight, values_from = score, names_prefix = "bws_") |> 
  relocate(GID_0, NAME_0, GID_1, NAME_1, bws_n, bws_dom, bws_ind, bws_irr, bws_liv, bws_tot)

## Don't try and deal with disputed areas
# |> 
#   mutate(GID_1 = case_when(
#     GID_1 == "IND.14_1" ~ "Z01.14_1",
#     GID_1 == "CHN.28_1" ~ "Z02.28_1",
#     GID_1 == "CHN.29_1" ~ "Z03.29_1",
#     GID_1 == "IND.13_1" ~ "Z04.13_1",
#     GID_1 == "IND.35_1" ~ "Z05.35_1",
#     GID_1 == "PAK.1_1" ~ "Z06.1_1",
#     GID_1 == "PAK.6_1" ~ "Z06.6_1",
#     GID_1 == "IND.3_1" ~ "Z07.3_1",
#     .default = GID_1
#   ))
# 
# bws_adm1 <- bws_adm1 |> 
#   bind_rows(
#     filter(bws_adm1, GID_1 == "Z02.28_1") |> mutate(GID_1 = "Z03.28_1"),
#     filter(bws_adm1, GID_1 == "Z03.29_1") |> mutate(GID_1 = "Z08.29_1"),
#     filter(bws_adm1, GID_1 == "Z04.13_1") |> mutate(GID_1 = "Z09.13_1"),
#     filter(bws_adm1, GID_1 == "Z05.35_1") |> mutate(GID_1 = "Z09.35_1")
#   )
```

Need to join RFS admin areas (ESRI) and WRI admin areas (GADM). This is made more difficult because ESRI layers and GADM admin layers don't have a single unified identifier. To create an ID that can match to WRI data, we'll need to do a spatial join to find the GADM GID_1 ID (i.e. state/province) that best matches the ESRI boundaries. Findest the largest overlap between GADM (v3.6) and ESRI layers.

```{r}
## There are 77 cases where there is no matching GID_1 ID (either because there is no spatial match or no GID_1 ID exists)
## Most of these are in the Marshall Islands
sf::sf_use_s2(FALSE)
spatial_join <- adm1 |> 
  sf::st_join(GADM_adm1_sf |> 
                select(GID_0, GID_1), largest = TRUE)

## Of those that have a GID_1 ID
## 74 don't find any match within the BWS data
spatial_join_bws <- spatial_join |>
  ## Don't need GID_0 because it's implicit in GID_1
  select(COUNTRY, ISO_3DIGIT, GID_0, NAME, GID_1) |> 
  left_join(bws_adm1 |> 
              select(GID_1, NAME_1, starts_with("bws")),
            by = c("GID_1"))
```

```{r}
## Could any of these 151 cases be matched with string distance techniques?
unmatched <- spatial_join_bws |> 
  filter(is.na(NAME_1)) |> 
  st_drop_geometry() |> 
  select(ISO_3DIGIT, GID_0, NAME_0 = NAME)

spatial_namejoin_bws_osa1 <- unmatched |> 
  mutate(NAME_clean = stringr::str_remove_all(stringi::stri_trans_general(NAME_0, "Latin-ASCII"), "`|'|-| "),
         ## Some areas how no matching GID
         GID_combo = case_when(is.na(GID_0) ~ ISO_3DIGIT,
                               TRUE ~ GID_0)) |> 
  fuzzyjoin::stringdist_left_join(bws_adm1 |> 
                                    mutate(NAME_clean = stringr::str_remove_all(stringi::stri_trans_general(NAME_1, "Latin-ASCII"), "`|'|-| ")) |> 
                                    select(GID_0, NAME_clean, starts_with("bws")),
                                  by = c("GID_combo" = "GID_0",
                                         "NAME_clean"), max_dist = 1)

spatial_namejoin_bws_soundex <- unmatched |> 
  mutate(NAME_clean = stringr::str_remove_all(stringi::stri_trans_general(NAME_0, "Latin-ASCII"), "`|'|-| "),
         ## Some areas how no matching GID
         GID_combo = case_when(is.na(GID_0) ~ ISO_3DIGIT,
                               TRUE ~ GID_0)) |> 
  fuzzyjoin::stringdist_left_join(bws_adm1 |> 
                                    mutate(NAME_clean = stringr::str_remove_all(stringi::stri_trans_general(NAME_1, "Latin-ASCII"), "`|'|-| ")),
                                  by = c("GID_combo" = "GID_0",
                                         "NAME_clean"),
                                  method = "soundex")

## No successful matches. We have to accept that these are just NAs
```

```{r}
## Now that we've joined this other data, we can use create the admin1 aggregates
adm1_joined <- spatial_join_bws |> 
  select(COUNTRY, ISO_3DIGIT, GID_0, NAME, GID_1, starts_with("bws")) |> 
  ## Make a rowID that can be used to join data back
  mutate(rowID = 1:n())

son_water_adm1 <- adm1_joined |> 
  # water depletion (wdp)
  left_join(exact_extract(list_wa$wdp_n, adm1_joined, fun = "mean", append_cols = "rowID", progress = TRUE) |> rename(wdp_n = mean), by = "rowID") |>
  left_join(aggregate_to_adm(list_wa$wdp_n, wdp_weights$dom, wdp_dom, adm1_joined, "rowID"), by = "rowID") |>
  left_join(aggregate_to_adm(list_wa$wdp_n, wdp_weights$ind, wdp_ind, adm1_joined, "rowID"), by = "rowID") |>
  left_join(aggregate_to_adm(list_wa$wdp_n, wdp_weights$irr, wdp_irr, adm1_joined, "rowID"), by = "rowID") |>
  left_join(aggregate_to_adm(list_wa$wdp_n, wdp_weights$liv, wdp_liv, adm1_joined, "rowID"), by = "rowID") |>
  left_join(aggregate_to_adm(list_wa$wdp_n, wdp_weights$tot, wdp_tot, adm1_joined, "rowID"), by = "rowID") |>
  # blue water scarcity (wsb)
  left_join(exact_extract(list_wa$wsb_n, adm1_joined, fun = "mean", append_cols = "rowID", progress = TRUE) |> rename(wsb_n = mean), by = "rowID") |>
  left_join(aggregate_to_adm(list_wa$wsb_n, wsb_weights$agr, wsb_agr, adm1_joined, "rowID"), by = "rowID") |>
  left_join(aggregate_to_adm(list_wa$wsb_n, wsb_weights$dom, wsb_dom, adm1_joined, "rowID"), by = "rowID") |>
  left_join(aggregate_to_adm(list_wa$wsb_n, wsb_weights$ind, wsb_ind, adm1_joined, "rowID"), by = "rowID") |>
  left_join(aggregate_to_adm(list_wa$wsb_n, wsb_weights$tot, wsb_tot, adm1_joined, "rowID"), by = "rowID") |> 
  # coastal eutrophication potential (cep)
  left_join(exact_extract(list_wp$cep_n, adm1_joined, fun = "mean", append_cols = "rowID", progress = TRUE) |> rename(cep_n = mean), by = "rowID") |>
  left_join(aggregate_to_adm(list_wp$cep_n, wp_weights, cep_agr, adm1_joined, "rowID"), by = "rowID") |>
  left_join(aggregate_to_adm(list_wp$cep_n, np_weights, cep_np, adm1_joined, "rowID"), by = "rowID")|>
  # nitrate-nitrite concentration (nox)
  left_join(exact_extract(list_wp$nox_n, adm1_joined, fun = "mean", append_cols = "rowID", progress = TRUE) |> rename(nox_n = mean), by = "rowID") |>
  left_join(aggregate_to_adm(list_wp$nox_n, wp_weights, nox_agr, adm1_joined, "rowID"), by = "rowID") |>
  left_join(aggregate_to_adm(list_wp$nox_n, np_weights, nox_np, adm1_joined, "rowID"), by = "rowID") |>
  # periphyton growth potential (pgp)
  left_join(exact_extract(list_wp$pgp_n, adm1_joined, fun = "mean", append_cols = "rowID", progress = TRUE) |> rename(pgp_n = mean), by = "rowID") |>
  left_join(aggregate_to_adm(list_wp$pgp_n, wp_weights, pgp_agr, adm1_joined, "rowID"), by = "rowID") |>
  left_join(aggregate_to_adm(list_wp$pgp_n, np_weights, pgp_np, adm1_joined, "rowID"), by = "rowID")|>
  
  mutate(across(where(is.numeric), ~ round(.x, 1))) |> 
  rowwise() |> 
  mutate(
    wa_max_n = max(c(bws_n, wdp_n, wsb_n), na.rm = TRUE),
    wa_max_dom = max(c(bws_dom, wdp_dom, wsb_dom), na.rm = TRUE),
    wa_max_ind = max(c(bws_ind, wdp_ind, wsb_ind), na.rm = TRUE),
    wa_max_irr = max(c(bws_irr, wdp_irr, wsb_agr), na.rm = TRUE),
    wa_max_liv = max(c(bws_liv, wdp_liv, wsb_n), na.rm = TRUE),
    wa_max_tot = max(c(bws_tot, wdp_tot, wsb_tot), na.rm = TRUE),
    wp_max_n = max(c(cep_n, nox_n, pgp_n), na.rm = TRUE),
    wp_max_agr = max(c(cep_agr, nox_agr, pgp_agr), na.rm = TRUE),
    wp_max_np = max(c(cep_np, nox_np, pgp_np), na.rm = TRUE)
  ) |> 
  ungroup() |> 
  relocate(wa_max_n:wa_max_tot, .before = bws_n) |>
  relocate(wp_max_n:wp_max_np, .before = cep_n) |> 
  relocate(geometry, .after = last_col()) |> 
  mutate(across(where(is.numeric), ~ if_else(between(.x, 1, 5), .x, NA)))
```

```{r}
st_drop_geometry(son_water_adm1) |> 
  skim()
```

## 6. Export

```{r}
key_descriptions <- tribble(
  ~Key,    ~Description,     
  "wa_max",    "Maximum value between water availability layers: bws, wdp, wsb", 
  "wp_max",  "Maximum value between water pollution layers: cep, nox, pgp", 
  "bws","Baseline Water Stress",     
  "wdp","Water depletion",      
  "wsb","Blue Water Scarcity",
  "cep","Coastal Eutrophication Potential",
  "nox","Nitrate-Nitrite Concentration",
  "pgp","Periphyton Growth Potential",
  "","",
  "","",
  "Suffix","Description",
  "_n","Aggregation at administrative level, without weighting",
  "_dom","Aggregation at administrative level weighted by domestic water demand*",
  "_ind","Aggregation at administrative level weighted by industrial water demand*",
  "_irr","Aggregation at administrative level weighted by irrigation water demand*",
  "_liv","Aggregation at administrative level weighted by livestock water demand*",
  "_tot","Aggregation at administrative level weighted by total water demand*",
  "_agr","Aggregation at administrative level weighted by cropland extent**",
  "","",
  "", "* Data for 'water demand' differs depending on the indicator aggregated",
  "bws","Water withdrawal from PCR-GLOBWB 2",
  "wdp","Water consumption from WaterGAP 3",
  "wsb","Blue water footprint from Water Footprint Network",
  "", "** For the water pollution indicators we used cropland extend from Potapov et al. 2021 (https://www.nature.com/articles/s43016-021-00429-z)"
)
```

```{r}
write.xlsx(list(
  "Help" = key_descriptions,
  "National level" = st_drop_geometry(son_water_adm0) |> 
    mutate(notes = if_else(GID_0 %in% c("RUS", "CAN", "CHN", "USA", "BRA", "AUS", "IND", "ARG", "KAZ", "DZA"),
                           "Not recommended to be used. Average values for such large country are not representative. Use values from sub-national level instead.", NA)), 
  "Sub-national level" = st_drop_geometry(son_water_adm1)
), file = here("/outputs", "adm_level", "sbtn_son_water_v1_1_adm_level.xlsx"), rowNames = FALSE)
```

*National level*

```{r}
saveRDS(son_water_adm0, here::here("/outputs", "adm_level", "sbtn_son_water_v1_1_adm0.rds"))
```

```{r}
# st_write(son_water_adm0, here("outputs", "adm_level", "sbtn_son_water_v1_1_adm0.shp"), layer_options = "ENCODING=UTF-8", delete_layer = TRUE)
```

*Sub-national level*

```{r}
saveRDS(son_water_adm1, here::here("/outputs", "adm_level", "sbtn_son_water_v1_1_adm1.rds"))
```

```{r}
# st_write(son_water_adm1, here("outputs", "adm_level", "sbtn_son_water_v1_1_adm1.shp"), layer_options = "ENCODING=UTF-8", delete_layer = TRUE)
```

## 7. View outputs

```{r}
son_water_adm0 <- readRDS(here::here("/outputs", "adm_level", "sbtn_son_water_v1_1_adm0.rds"))
son_water_adm1 <- readRDS(here::here("/outputs", "adm_level", "sbtn_son_water_v1_1_adm1.rds"))
```

```{r}
mapview_custom <- function(data, layer_name, var){
  mapview(
    data,
    layer.name = layer_name,
    zcol = var,
    col.regions = colorRampPalette(c("#FFFF99", "#FFE600", "#FF9900", "#FF1900", "#990000")),
    at = c(1, 1.8, 2.6, 3.4, 4.2, 5),
    lwd = 0.1
  )
}
```

*National level*

```{r}
son_water_adm0_simplified <- son_water_adm0 |> 
  st_transform("EPSG:8857") |> 
  st_simplify(dTolerance = 1000) |> 
  st_transform("EPSG:4326")
```

```{r}
mapview_custom(son_water_adm0_simplified, "Water Availability (Max)", "wa_max_tot")
```

```{r}
mapview_custom(son_water_adm0_simplified, "Water Pollution (Max)", "wp_max_agr")
```

*Sub-national level*

```{r}
son_water_adm1_simplified <- son_water_adm1 |> 
  st_transform("EPSG:8857") |> 
  st_simplify(dTolerance = 1000) |> 
  st_transform("EPSG:4326")
```

```{r}
mapview_custom(son_water_adm1_simplified, "Water Availability (Max)", "wa_max_tot")
```

```{r}
mapview_custom(son_water_adm1_simplified, "Water Pollution (Max)", "wp_max_agr")
```
