---
title:  "SBTN State of Nature Water unified layers (v2.1)"
author: "Liam D. Bailey"
lightbox: true
format:
    html:
      toc-title: Table of contents
      toc: true
      toc-depth: 2
      toc-location: left
      toc-expand: true
      fontsize: "10pt"
      fontcolor: "black"
      margin-left: "15pt"
      margin-right: "15pt"
      number-sections: true
      highlight-style: github
      cap-location: bottom
      page-layout: full
      embed-resources: true
      self-contained-math: true
      theme:
        - journal
execute: 
  eval: true
  warning: false
params:
  sharepoint: "C:\\Users\\liam.bailey\\WWF-Deutschland\\FB-WuM-Daten - Water & Biodiversity Stewardship"
  overwrite: FALSE
---

```{r}
#| echo: false
#| include: false

## SETUP
if (!require("pacman")) install.packages("pacman")
pacman::p_load(here, dplyr, magrittr, sf, terra, stars, exactextractr, skimr, readr, tidyr, openxlsx, mapview, ggplot2)
sf::sf_use_s2(FALSE) # due to issues in HydroBASINS geometries
```

## Introduction

::: callout-note
-   Unified layer v2.0 replaces Blue Water Scarcity (Mekonnen & Hoekstra) and [Freshwater Quantity Targets v1.0 (Hogeboom et al. 2024)](https://doi.org/10.4121/089afa18-bbd5-493e-8a6c-39e1ca7b2a59.v1).
-   Unified layer v2.1 uses newly released [Freshwater Quantity Targets v2.0 (Hogeboom et al. 2024)](https://doi.org/10.4121/089afa18-bbd5-493e-8a6c-39e1ca7b2a59.v2).
:::

## 1. Load data

### 1.1. Water Availability

#### 1.1.1. Baseline Water Stress [Kuzma et al. 2023](https://www.wri.org/data/aqueduct-global-maps-40-data)

```{r}
bws <- st_read(file.path(params$sharepoint, "10_RiskSuite", "RFS_New_Folders", "05_Data_and_Analysis", "01_Data", "Aqueduct_WRI", "v4.0", "GDB", "Aq40_Y2023D07M05.gdb"),
               layer = "baseline_annual")
```

#### 1.1.2. Water Depletion [Brauman et al. 2016](http://www.earthstat.org/water-depletion-watergap3-basins/)

```{r}
wdp <- rast(file.path(params$sharepoint, "10_RiskSuite", "RFS_New_Folders", "05_Data_and_Analysis", "01_Data", "WaterAvailability_WaterDepletion_WaterGap3", "WaterDepletion", "WaterDepletionCat_WG3.tif"))
```

#### 1.1.3. Freshwater Quantity Targets v2.0 [Hogeboom et al. 2024](https://doi.org/10.4121/089afa18-bbd5-493e-8a6c-39e1ca7b2a59.v2)

Also called 'Blue Water Reduction Target' in app recommended by SBTN (https://www.waterfootprintassessmenttool.org/hbasins/sustainability/map/blue?b=sbtn)

```{r}
fqt <- read_sf(file.path(params$sharepoint, "10_RiskSuite", "RFS_New_Folders", "05_Data_and_Analysis", "01_Data", "WaterAvailability_FreshwaterQuantityTargets_Hogeboom24", "v2.0", "FQT_v2.gpkg"))
```

### 1.2. Water Pollution

#### 1.2.1. Coastal Eutrophication Potential [Kuzma et al. 2023](https://www.wri.org/data/aqueduct-global-maps-40-data)

Comes from Aqueduct 4.0 data (but unchanged since Aqueduct 3.0)

```{r}
cep <- bws
```

#### 1.2.2. Nitrate-Nitrite Concentration [Damania et al. 2019](https://wbwaterdata.org/dataset/global-nitrate-nitrite-surface-water)

```{r}
nox_csv <- read_csv(file.path(params$sharepoint, "10_RiskSuite", "RFS_New_Folders", "05_Data_and_Analysis", "01_Data", "WaterQuality_WorldBank", "WorldBank_ML_Quality", "WorldBank_ML_Quality", "noxn_preds_05_16_19.csv"))
```

*Spatial grid*

```{r}
sp_grid <- read_csv(file.path(params$sharepoint, "10_RiskSuite", "RFS_New_Folders", "05_Data_and_Analysis", "01_Data", "WaterQuality_WorldBank", "WorldBank_ML_Quality", "WorldBank_ML_Quality", "ML_data_gridcell_match.csv"))
```

#### 1.2.3. Periphyton Growth Potential [McDowell et al. 2020](https://figshare.com/s/e44fe7319585183d888f)

*Total Nitrogen Concentration*

```{r}
tnc <- rast(file.path(params$sharepoint, "10_RiskSuite", "RFS_New_Folders", "05_Data_and_Analysis", "01_Data", "WaterQuality_FreshwaterNutrientPollution_McDowell20", "data_from_author", "WQ_Variables_Extraction", "TN_Conc.tif"))
```

*Total Phosphorus Concentration*

```{r}
tpc <- rast(file.path(params$sharepoint, "10_RiskSuite", "RFS_New_Folders", "05_Data_and_Analysis", "01_Data", "WaterQuality_FreshwaterNutrientPollution_McDowell20", "data_from_author", "WQ_Variables_Extraction", "TN_Conc.tif"))
                      "TP_Conc.tif"))
```

### 1.3. Aggregation unit

#### 1.3.1. WMO Basins and Sub-Basins [GRDC 2020](https://panda.maps.arcgis.com/home/item.html?id=be4b6f13121b4670ad8f006bc1908e14)

*To add name of basins*

```{r}
wmo_basins <- st_read(file.path(params$sharepoint, "10_RiskSuite", "RFS_New_Folders", "05_Data_and_Analysis", "01_Data", "Basins_WMOBasin_GDRC", "wmobb_basins.shp")) |> 
  select(WMOBB_id = WMOBB, WMOBB_name = WMOBBName, region = REGNAME)
```

#### 1.3.2. HydroSHEDS HydroBASINS Level 6 [Lehner & Grill 2013](https://www.hydrosheds.org/products/hydrobasins)

```{r}
if (params$overwrite | !file.exists(here("./v2.1/outputs", "h6.rds"))){
  h6 <- st_read(file.path(params$sharepoint,
                          # "10_RiskSuite", "RFS_New_Folders", "05_Data_and_Analysis", "01_Data", "Basins_HydroSHEDS_McGill", "v1.0", "HydroBASINS",
                          "hybas_lev06_v1c.gpkg")) |> 
    select(HYBAS_ID) |>
    mutate(HYBAS_ID = as.factor(HYBAS_ID)) |>
    ## Link WMO basin with Hydroshed 6 that has the largest overlap
    st_join(wmo_basins, largest = TRUE)
  
  write_rds(h6, here("./v2.1/outputs", "h6.rds")) 
} else {
  h6 <- read_rds(here("./v2.1/outputs", "h6.rds"))
}
```

## 2. Clean and Classify data to 1-5 values

### 2.1. Water Availability

#### 2.1.1. Baseline Water Stress

```{r}
bws_n <- bws |>
  mutate(
    bws_raw = na_if(bws_raw, -9999),
    bws_raw = na_if(bws_raw, 9999),
    bws_n = case_when(
      bws_label == "Low (<10%)" ~ 1,
      bws_label == "Low - Medium (10-20%)" ~ 2,
      bws_label == "Medium - High (20-40%)" ~ 3,
      bws_label == "High (40-80%)" ~ 4,
      bws_label %in% c("Extremely High (>80%)", "Arid and Low Water Use") ~ 5
    )
  ) |>
  select(bws_raw, bws_n, bws_label, geometry = Shape)
```

#### 2.1.2. Water Depletion

```{r}
wdp_reclass <- matrix(
  c(
    0,NA,
    1,1, # <=5% annual depletion
    2,2, # >5–25% annual depletion
    5,3, # >25–75% annual depletion, Dry-Year
    6,4, # >25–75% annual depletion, Seasonal
    7,5, # >75–100% annual depletion
    8,5  # >100% annual depletion
  ),
  ncol=2, byrow=TRUE
)
```

```{r}
wdp_n <- wdp |>
  classify(wdp_reclass) |> 
  exact_extract(h6, fun = c("mean", "median", "quantile", "majority", "coefficient_of_variation"), quantiles = 0.75,
                append_cols = "HYBAS_ID", progress = TRUE) 

# skim(wdp_n)
```

```{r}
wdp_n <- wdp_n |>
  mutate(
    wdp_raw = median,
    wdp_n = round(median),
    wdp_label = case_when(
      wdp_n == 1 ~ "<=5% annual depletion",
      wdp_n == 2 ~ ">5–25% annual depletion",
      wdp_n == 3 ~ ">25–75% annual depletion, Dry-Year",
      wdp_n == 4 ~ ">25–75% annual depletion, Seasonal",
      wdp_n == 5 ~ ">75% annual depletion"
    )
  ) |>
  select(HYBAS_ID, wdp_raw:wdp_label)
```

#### 2.1.3. Freshwater Quantity Targets

```{r}
fqt_n <- fqt |>
  mutate(
    fqt_raw = Target_annual,
    fqt_n = case_when(
      fqt_raw < 0.05 ~ 1,
      fqt_raw < 0.25 ~ 2,
      fqt_raw < 0.5 ~ 3,
      fqt_raw < 0.8 ~ 4,
      fqt_raw >= 0.8 ~ 5
    ),
    fqt_label = case_when(
      fqt_n == 1 ~ "Low (<5%)",
      fqt_n == 2 ~ "Moderate (5-25%)",
      fqt_n == 3 ~ "Significant (25-50%)",
      fqt_n == 4 ~ "Severe (50-80%)",
      fqt_n == 5 ~ "Extreme (>=80%)"
    )
  ) |>
  select(fqt_raw, fqt_n, fqt_label)
```

#### 2.1.4. Join datasets and Export

```{r}
if (params$overwrite | !file.exists(here("./v2.1/outputs", "wa_n.rds"))) {
  wa_n <- h6 |>
    st_join(bws_n, largest = TRUE) |>
    left_join(wdp_n, by = "HYBAS_ID") |>
    st_join(fqt_n, largest = TRUE) 
  write_rds(wa_n, here("./v2.1/outputs", "wa_n.rds"))
} else {
  wa_n <- read_rds(here("./v2.1/outputs", "wa_n.rds"))
}
```

### 2.2. Water Pollution

#### 2.2.1. Coastal Eutrophication Potential

```{r}
cep_n <- cep |>
  mutate(
    cep_raw = na_if(cep_raw, -9999),
    cep_raw = na_if(cep_raw, 9999),
    cep_n = case_when(
      cep_label == "Low (<-5)" ~ 1,
      cep_label == "Low - Medium (-5 to 0)" ~ 2,
      cep_label == "Medium - High (0 to 1)" ~ 3,
      cep_label == "High (1 to 5)" ~ 4,
      cep_label == "Extremely High (>5)" ~ 5
    )
  ) |>
  select(cep_raw, cep_n, cep_label, geometry = Shape)
```

#### 2.2.2. Nitrate-Nitrite Concentration

*Get mean predicted values between 2006 and 2010, then rasterize*

```{r}
nox <- nox_csv |>
  filter(year >= 2006) |>
  group_by(OBJECTID) |>
  summarise(value = mean(preds)) |>
  right_join(sp_grid, by = "OBJECTID") |>
  select(-OBJECTID) |>
  st_as_sf(coords = c("Lon", "Lat"), crs = 4326) |>
  st_rasterize(dx = 0.5, dy = 0.5) |>
  rast()
```

```{r}
nox_n <- exact_extract(nox, h6, fun = c("mean", "median", "quantile", "majority", "coefficient_of_variation"), quantiles = 0.75,
                       append_cols = "HYBAS_ID", progress = TRUE)

# skim(nox_n)
```

```{r}
nox_n <- nox_n |>
  mutate(
    nox_raw = median,
    nox_n = case_when(
      median <= 0.4 ~ 1,
      median > 0.4 & median <= 0.8 ~ 2,
      median > 0.8 & median <= 1.2 ~ 3,
      median > 1.2 & median <= 1.6 ~ 4,
      median > 1.6 ~ 5
    ),
    nox_label = case_when(
      nox_n == 1 ~ "Very Low Concentration (<=0.4)",
      nox_n == 2 ~ "Low Concentration (>0.4-0.8)",
      nox_n == 3 ~ "Moderate Concentration (>0.8-1.2)",
      nox_n == 4 ~ "High Concentration (>1.2-1.6)",
      nox_n == 5 ~ "Very High Concentration (>1.6)"
    )
  ) |>
  select(HYBAS_ID, nox_raw:nox_label)
```

#### 2.2.3. Periphyton Growth Potential

```{r}
tnc_n <- exact_extract(tnc, h6, fun = c("mean", "median", "quantile", "majority", "coefficient_of_variation"), quantiles = 0.75,
                       append_cols = "HYBAS_ID", progress = TRUE)

# skim(tnc_n)
```

```{r}
tpc_n <- exact_extract(tpc, h6, fun = c("mean", "median", "quantile", "majority", "coefficient_of_variation"), quantiles = 0.75,
                       append_cols = "HYBAS_ID", progress = TRUE)

# skim(tpc_n)
```

```{r}
pgp_n <- tnc_n |>
  select(HYBAS_ID, tnc_raw = median) |>
  left_join(
    select(tpc_n, HYBAS_ID, tpc_raw = median),
    by = "HYBAS_ID" 
  ) |>
  mutate(
    ratio_n_p = tnc_raw/tpc_raw,
    limiting = if_else(ratio_n_p < 7, "N-limited", "P-limited"),
    pgp_n = case_when(
      limiting == "N-limited" & tnc_raw <= 0.4 ~ 1,
      limiting == "N-limited" & tnc_raw > 0.4 & tnc_raw <= 0.8 ~ 2,
      limiting == "N-limited" & tnc_raw > 0.8 & tnc_raw <= 1.2 ~ 3,
      limiting == "N-limited" & tnc_raw > 1.2 & tnc_raw <= 1.6 ~ 4,
      limiting == "N-limited" & tnc_raw > 1.6 ~ 5,
      
      limiting == "P-limited" & tpc_raw <= 0.023 ~ 1,
      limiting == "P-limited" & tpc_raw > 0.023 & tpc_raw <= 0.046 ~ 2,
      limiting == "P-limited" & tpc_raw > 0.046 & tpc_raw <= 0.100  ~ 3,
      limiting == "P-limited" & tpc_raw > 0.100 & tpc_raw <= 0.150 ~ 4,
      limiting == "P-limited" & tpc_raw > 0.150 ~ 5
    ),
    pgp_label = case_when(
      limiting == "N-limited" & pgp_n == 1 ~ "N-limited growth acceptable",
      limiting == "N-limited" & pgp_n == 2 ~ "N-limited growth acceptable",
      limiting == "N-limited" & pgp_n == 3 ~ "N-limited growth undesirable",
      limiting == "N-limited" & pgp_n == 4 ~ "N-limited growth undesirable",
      limiting == "N-limited" & pgp_n == 5 ~ "N-limited growth undesirable",
      
      limiting == "P-limited" & pgp_n == 1 ~ "P-limited growth acceptable",
      limiting == "P-limited" & pgp_n == 2 ~ "P-limited growth acceptable",
      limiting == "P-limited" & pgp_n == 3 ~ "P-limited growth undesirable",
      limiting == "P-limited" & pgp_n == 4 ~ "P-limited growth undesirable",
      limiting == "P-limited" & pgp_n == 5 ~ "P-limited growth undesirable"
    )
  ) |>
  select(HYBAS_ID, tnc_raw, tpc_raw, ratio_n_p, limiting, pgp_n, pgp_label)
```

#### 2.2.4. Join datasets and Export

```{r}
if(params$overwrite | !file.exists(here("./v2.1/outputs", "wp_n.rds"))){
  wp_n <- h6 |>
    st_join(cep_n, largest = TRUE) |>
    left_join(nox_n, by = "HYBAS_ID") |>
    left_join(pgp_n, by = "HYBAS_ID")
  write_rds(wp_n, here("./v2.1/outputs", "wp_n.rds"))
} else {
  wp_n <- read_rds(here("./v2.1/outputs", "wp_n.rds"))
}
```

## 3. Identify the value for the unified layer

### 3.1. Max of Water Availability

```{r}
son_wa <- wa_n |>
  rowwise() |>
  mutate(
    wa_max = max(c(bws_n, wdp_n, fqt_n), na.rm = TRUE),
    wa_max = if_else(between(wa_max, 1, 5), wa_max, NA),
    .after = region
  ) |>
  ungroup()
```

### 3.2. Max of Water Pollution

```{r}
son_wp <- wp_n |>
  rowwise() |>
  mutate(
    wp_max = max(c(cep_n, nox_n, pgp_n), na.rm = TRUE),
    wp_max = if_else(between(wp_max, 1, 5), wp_max, NA),
    .after = region
  ) |>
  ungroup()
```

## 4. Join final output

```{r}
son_water <- son_wa |>
  left_join(
    son_wp |> select(-c(WMOBB_id:region)) |> st_drop_geometry(),
    by = "HYBAS_ID"
  ) |>
  relocate(geom, .after = last_col())

# st_drop_geometry(son_water) |> skim()
```

```{r}
son_water |> 
  st_transform("EPSG:8857") |> 
  st_simplify(dTolerance = 10000) |> 
  st_transform(st_crs(son_water)) |> 
  mutate(across(c(wa_max, wp_max), ~ as.factor(.x))) |> 
  mapview(
    layer.name = list("Max Water Availability", "Max Water Pollution"),
    zcol = c("wa_max", "wp_max"),
    col.regions = c("#FFFF99", "#FFE600", "#FF9900", "#FF1900", "#990000"),
    lwd = 0.2,
    alpha.regions = 0.9
  )
```

## 5. Sanity check

Read in old data from last version and check where they differ.

```{r}
v2.0 <- read_excel(here("v2.0/outputs/sbtn_son_water_v2.0.xlsx"))
```

```{r}
compare_wa <- son_water |> 
  st_drop_geometry() |> 
  left_join(v2.0 |> 
              select(HYBAS_ID, wa_max_2.0 = wa_max, fqt_raw_2.0 = fqt_raw, bws_raw_2.0 = bws_raw, wdp_raw_2.0 = wdp_raw),
            by = "HYBAS_ID") |> 
  mutate(wa_diff = wa_max - wa_max_2.0)
```

```{r}
## We expect some small changes due to new Hogeboom data
hist(compare_wa$wa_diff)
```

```{r}
hist(round(compare_wa$fqt_raw - compare_wa$fqt_raw_2.0, 5))
```

```{r}
hist(round(compare_wa$bws_raw - compare_wa$bws_raw_2.0, 5))
```

```{r}
hist(round(compare_wa$wdp_raw - compare_wa$wdp_raw_2.0, 5))
```

```{r}
compare_wp <- son_water |> 
  st_drop_geometry() |> 
  left_join(v2.0 |> 
              select(HYBAS_ID, wp_max_2.0 = wp_max, cep_raw_2.0 = cep_raw, nox_raw_2.0 = nox_raw, tnc_raw_2.0 = tnc_raw, tpc_raw_2.0 = tpc_raw),
            by = "HYBAS_ID") |> 
  mutate(wp_diff = wp_max - wp_max_2.0) |> 
  arrange(desc(wp_diff)) |> 
  select(HYBAS_ID, starts_with("wp_"), cep_raw, cep_raw_2.0, nox_raw, nox_raw_2.0, tnc_raw, tnc_raw_2.0, tpc_raw, tpc_raw_2.0)

compare_wp
```

```{r}
hist(round(compare_wp$cep_raw - compare_wp$cep_raw_2.0, 5))
```

```{r}
hist(round(compare_wp$nox_raw - compare_wp$nox_raw_2.0, 5))
```

```{r}
hist(round(compare_wp$tnc_raw - compare_wp$tnc_raw_2.0, 5))
```

```{r}
hist(round(compare_wp$tpc_raw - compare_wp$tpc_raw_2.0, 5))
```

## 6. Export

```{r}
write_rds(son_water, here("./v2.1/outputs", "sbtn_son_water_v2.1.rds"))
```

```{r}
st_write(son_water, here("./v2.1/outputs", "sbtn_son_water_v2.1.shp"), layer_options = "ENCODING=UTF-8", delete_layer = TRUE)
```

```{r}
write.xlsx(list(
  "SBTN SoN Water" = st_drop_geometry(son_water)
), file = here("./v2.1/outputs", "sbtn_son_water_v2.1.xlsx"), rowNames = FALSE)
```
