---
title: "SBTN State of Nature Water unified layers"
subtitle: "water availability and water pollution - National and Sub-national level, weighted by sector"
author: "Liam D. Bailey"
date: "July 2, 2024"
engine: knitr
format:
  html:
    toc: true
execute:
  warning: false
params:
  sharepoint: "/Users/liam/Downloads"
  version: "v2.1"
  overwrite: FALSE
---

## 1. Setup

Load required packages

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(here, dplyr, readr, readxl, sf, tidyr, terra, fasterize, purrr, exactextractr, skimr, mapview, openxlsx)
mem.maxVSize(30000)
```

```{r}
dir.create(here(params$version, "outputs", "temp_raster"))
```

## 2. Load data

```{r}
son_water <- read_rds(here(params$version, "outputs", paste0("sbtn_son_water_", params$version, ".rds")))
```

```{r}
aqueduct_adm0 <- read_excel(
  file.path(
    params$sharepoint,
    # "10_RiskSuite", "RFS_New_Folders", "05_Data_and_Analysis", "01_Data", "Aqueduct_WRI", "v4.0", "GDB",
    "Aqueduct40_rankings_download_Y2023M07D05.xlsx"
  ),
  sheet = "country_baseline")
```

```{r}
aqueduct_adm1 <- read_excel(
  file.path(
    params$sharepoint,
    # "10_RiskSuite", "RFS_New_Folders", "05_Data_and_Analysis", "01_Data", "Aqueduct_WRI", "v4.0", "GDB",
    "Aqueduct40_rankings_download_Y2023M07D05.xlsx"
  ), sheet = "province_baseline")
```

*National level administrative division*

Here we are using our RFS assessment units (i.e. separate Hawaii and Azores). For consistency, we will also create a copy that also uses GADM 4.1 to upload onto Zenodo.

```{r}
adm0 <- read_sf(
  file.path(
    params$sharepoint,
    # "10_RiskSuite", "RFS_New_Folders", "05_Data_and_Analysis", "02_Analysis", "RFSAssessmentUnits", "output", "final",
    "RFSlocations_EPSG4326.gpkg"
  )) |> 
  filter(SOURCE %in% c("World Bank", "HydroBASIN7")) |> 
  rename(LOCATION_ID = LOCATION)
```

*Sub-national level administrative division*

```{r}
adm1 <- st_read(
  file.path(
    params$sharepoint,
    # "10_RiskSuite/RFS_New_Folders/05_Data_and_Analysis/01_Data/Borders_WorldBank",
    "World Bank Official Boundaries - Admin 1.gpkg"
  ), quiet = TRUE
)
```

Read in translation table so we can match to Aqueduct data

```{r}
GADM_trans <- read.csv(
  file.path(
    params$sharepoint,
    "GADM_WB_translation.csv"
  )
)
```

## 3. Rasterize SoN water layers

```{r}
fasterize_son_water <- function(var){
  if (!file.exists(here(params$version, "outputs", "temp_raster", paste0(var, ".tif"))) | (file.exists(here(params$version, "outputs", "temp_raster", paste0("wdp_weight_", var, ".tif"))) & params$overwrite)){
    fasterize(son_water, raster(son_water, res = 0.01), field = var, fun = "max") |>
      terra::writeRaster(here(params$version, "outputs", "temp_raster", paste0(var, ".tif")), overwrite = TRUE)
  }
  return(NULL)
}

map(list("wdp_n", "fqt_n", "cep_n", "nox_n", "pgp_n"), fasterize_son_water)
```

```{r}
list_wa <- list(
  rast(here(params$version, "outputs", "temp_raster", "wdp_n.tif")),
  rast(here(params$version, "outputs", "temp_raster", "fqt_n.tif"))
) |> 
  set_names("wdp_n", "fqt_n")
```

```{r}
list_wp <- list(
  rast(here(params$version, "outputs", "temp_raster", "cep_n.tif")),
  rast(here(params$version, "outputs", "temp_raster", "nox_n.tif")),
  rast(here(params$version, "outputs", "temp_raster", "pgp_n.tif"))
) |> 
  set_names("cep_n", "nox_n", "pgp_n")
```

## 4. Prepare data to use as weights

### 4.1. WaterGap3 water consumption as weights for water depletion (provided by the authors)

```{r}
waterGap3_consumption <- read_csv(
  file.path(
    params$sharepoint,
    # "10_RiskSuite", "RFS_New_Folders", "05_Data_and_Analysis", "01_Data", "WaterAvailability_WaterDepletion_WaterGap3", "WaterConsumption",
    "SectorConsumption_m3_annual.csv"
  ),
  # here("inputs", "water_depletion", "SectorConsumption_m3_annual.csv"),
  col_names = c("subBas_ID", "domestic", "manufacturing", "electric", "irrigation", "livestock")
)

waterGap3_consumption <- read_sf(
  file.path(
    params$sharepoint,
    # "10_RiskSuite", "RFS_New_Folders", "05_Data_and_Analysis", "01_Data", "WaterAvailability_WaterDepletion_WaterGap3", "WaterGap3Shape",
    "WaterGap3_AllBasins.shp"
  )) |> 
  # here("inputs", "water_depletion", "WaterGap3Shape", "WaterGap3_AllBasins.shp")) |>
  dplyr::select(subBas_ID = ContUID) |>
  inner_join(waterGap3_consumption, by = "subBas_ID") |>
  mutate(
    industrial = manufacturing + electric,
    total = domestic + irrigation + livestock + industrial
  ) |>
  dplyr::select(
    dom = domestic,
    ind = industrial,
    irr = irrigation,
    liv = livestock,
    tot = total
  )

fasterize_vars <- function(var){
  if (!file.exists(here(params$version, "outputs", "temp_raster", paste0("wdp_weight_", var, ".tif"))) | (file.exists(here(params$version, "outputs", "temp_raster", paste0("wdp_weight_", var, ".tif"))) & params$overwrite)){
    fasterize(waterGap3_consumption, raster(waterGap3_consumption, res = 0.01), field = var, fun = "max") |>
      writeRaster(here(params$version, "outputs", "temp_raster", paste0("wdp_weight_", var, ".tif")), overwrite = TRUE) 
  }
  return(NULL)
}

map(list("dom", "ind", "irr", "liv", "tot"), fasterize_vars)
```

```{r}
wdp_weights <- list.files(here(), pattern = "wdp_weight_", recursive = TRUE, full.names = TRUE) |> 
  map(rast) |> 
  set_names("dom", "ind", "irr", "liv", "tot")
```

### 4.2. Water Footprint Network Blue water footprint as weights for blue water scarcity (provided by the authors)

```{r}
fqt_weights <- list(
  ## TODO: replace with H-BWF
  rast(
    file.path(
      params$sharepoint,
      # "10_RiskSuite/RFS_New_Folders/05_Data_and_Analysis/01_Data/WaterAvailability_BlueWaterFootprint_MekonnenHoekstra/WFbl_AgDomInd/WFbl_m3yr/",
      "wfbl_agr_m3y",
      "w001001.adf"
    )
  ),
  rast(
    file.path(
      params$sharepoint,
      # "10_RiskSuite/RFS_New_Folders/05_Data_and_Analysis/01_Data/WaterAvailability_BlueWaterFootprint_MekonnenHoekstra/WFbl_AgDomInd/WFbl_m3yr/",
      "wfbl_dom_m3y",
      "w001001.adf"
    )
  ),
  rast(
    file.path(
      params$sharepoint,
      # "10_RiskSuite/RFS_New_Folders/05_Data_and_Analysis/01_Data/WaterAvailability_BlueWaterFootprint_MekonnenHoekstra/WFbl_AgDomInd/WFbl_m3yr/",
      "wfbl_ind_m3y",
      "w001001.adf"
    )
  ),
  rast(
    file.path(
      params$sharepoint,
      # "10_RiskSuite/RFS_New_Folders/05_Data_and_Analysis/01_Data/WaterAvailability_BlueWaterScarcity_MekonnenHoekstra/WS_blue_monthly_rasters/WSbl_monthly_30m",
      "ws_avg",
      "w001001.adf"
    )
  )
) |> 
  set_names("agr", "dom", "ind", "tot")
```

### 4.3. Global Cropland Extent (2019) [Potapov et al. 2021](https://glad.umd.edu/dataset/croplands)

```{r}
wp_weights <- rast(
  file.path(
    params$sharepoint,
    # "10_RiskSuite/RFS_New_Folders/05_Data_and_Analysis/01_Data/Agriculture_CroplandExtent_Potapov21",
    "Global_cropland_3km_2019.tif"
  )
)
```

## 5. Spatial aggregation to administrative divisions

```{r}
aggregate_to_adm <- function(indicator, weight_by, key, to_adm, adm_id){
  exact_extract(indicator, to_adm, fun = "weighted_mean", weights = resample(weight_by, indicator), default_weight = 0, append_cols = adm_id, progress = TRUE) |> 
    rename({{key}} := weighted_mean)
} 
```

### 5.1. National level

```{r}
bws_adm0 <- aqueduct_adm0 |> 
  filter(indicator_name == "bws") |> 
  mutate(
    weight = if_else(weight == "One", "n", tolower(weight)),
    score = na_if(score, -9999),
    # rescale values 0-5 to 1-5 applying (new_max - new_min) * (x - min_old) / (max_old - min_old) + min_new
    score = (5-1) * ((score-0) / (5-0)) + 1 
  ) |> 
  select(LOCATION_ID = gid_0, weight, score) |> 
  pivot_wider(names_from = weight, values_from = score, names_prefix = "bws_") |> 
  relocate(LOCATION_ID, bws_n, bws_dom, bws_ind, bws_irr, bws_liv, bws_tot)
```

```{r}
bws_adm0_locationID <- adm0 |> 
  left_join(bws_adm0, by = "LOCATION_ID")

bws_adm0_sovereign <- adm0 |> 
  left_join(bws_adm0, by = c("SOVEREIGN" = "LOCATION_ID"))

bws_adm0_combo <- bws_adm0_locationID |> 
  filter(!is.na(bws_n)) |> 
  bind_rows(
    bws_adm0_sovereign |> 
      filter(!LOCATION_ID %in% bws_adm0_locationID$LOCATION_ID[!is.na(bws_adm0_locationID$bws_n)] & !is.na(bws_n))
  )

son_water_adm0_extract <- adm0 |>
  # baseline water stress (bws)
  left_join(bws_adm0_combo |> st_drop_geometry(), by = "LOCATION_ID") |> 
  # water depletion (wdp)
  left_join(exact_extract(list_wa$wdp_n, adm0, fun = "mean", append_cols = "LOCATION_ID", progress = TRUE) |> rename(wdp_n = mean), by = "LOCATION_ID") |>
  left_join(aggregate_to_adm(list_wa$wdp_n, wdp_weights$dom, wdp_dom, adm0, "LOCATION_ID"), by = "LOCATION_ID") |>
  left_join(aggregate_to_adm(list_wa$wdp_n, wdp_weights$ind, wdp_ind, adm0, "LOCATION_ID"), by = "LOCATION_ID") |>
  left_join(aggregate_to_adm(list_wa$wdp_n, wdp_weights$irr, wdp_irr, adm0, "LOCATION_ID"), by = "LOCATION_ID") |>
  left_join(aggregate_to_adm(list_wa$wdp_n, wdp_weights$liv, wdp_liv, adm0, "LOCATION_ID"), by = "LOCATION_ID") |>
  left_join(aggregate_to_adm(list_wa$wdp_n, wdp_weights$tot, wdp_tot, adm0, "LOCATION_ID"), by = "LOCATION_ID") |>
  # freshwater quantity target (fqt)
  left_join(exact_extract(list_wa$fqt_n, adm0, fun = "mean", append_cols = "LOCATION_ID", progress = TRUE) |> rename(fqt_n = mean), by = "LOCATION_ID") |>
  left_join(aggregate_to_adm(list_wa$fqt_n, fqt_weights$agr, fqt_agr, adm0, "LOCATION_ID"), by = "LOCATION_ID") |>
  left_join(aggregate_to_adm(list_wa$fqt_n, fqt_weights$dom, fqt_dom, adm0, "LOCATION_ID"), by = "LOCATION_ID") |>
  left_join(aggregate_to_adm(list_wa$fqt_n, fqt_weights$ind, fqt_ind, adm0, "LOCATION_ID"), by = "LOCATION_ID") |>
  left_join(aggregate_to_adm(list_wa$fqt_n, fqt_weights$tot, fqt_tot, adm0, "LOCATION_ID"), by = "LOCATION_ID") |> 
  # coastal eutrophication potential (cep)
  left_join(exact_extract(list_wp$cep_n, adm0, fun = "mean", append_cols = "LOCATION_ID", progress = TRUE) |> rename(cep_n = mean), by = "LOCATION_ID") |>
  left_join(aggregate_to_adm(list_wp$cep_n, wp_weights, cep_agr, adm0, "LOCATION_ID"), by = "LOCATION_ID") |>
  # nitrate-nitrite concentration (nox)
  left_join(exact_extract(list_wp$nox_n, adm0, fun = "mean", append_cols = "LOCATION_ID", progress = TRUE) |> rename(nox_n = mean), by = "LOCATION_ID") |>
  left_join(aggregate_to_adm(list_wp$nox_n, wp_weights, nox_agr, adm0, "LOCATION_ID"), by = "LOCATION_ID") |>
  # periphyton growth potential (pgp)
  left_join(exact_extract(list_wp$pgp_n, adm0, fun = "mean", append_cols = "LOCATION_ID", progress = TRUE) |> rename(pgp_n = mean), by = "LOCATION_ID") |>
  left_join(aggregate_to_adm(list_wp$pgp_n, wp_weights, pgp_agr, adm0, "LOCATION_ID"), by = "LOCATION_ID")

## Calculate max values for all
son_water_adm0 <- son_water_adm0_extract |> 
  mutate(across(where(is.numeric), ~ round(.x, 1))) |> 
  rowwise() |> 
  mutate(
    wa_max_n = max(c(bws_n, wdp_n, fqt_n), na.rm = TRUE),
    wa_max_dom = max(c(bws_dom, wdp_dom, fqt_dom), na.rm = TRUE),
    wa_max_ind = max(c(bws_ind, wdp_ind, fqt_ind), na.rm = TRUE),
    wa_max_irr = max(c(bws_irr, wdp_irr, fqt_agr), na.rm = TRUE),
    wa_max_liv = max(c(bws_liv, wdp_liv, fqt_n), na.rm = TRUE),
    wa_max_tot = max(c(bws_tot, wdp_tot, fqt_tot), na.rm = TRUE),
    wp_max_n = max(c(cep_n, nox_n, pgp_n), na.rm = TRUE),
    wp_max_agr = max(c(cep_agr, nox_agr, pgp_agr), na.rm = TRUE)
  ) |> 
  ungroup() |> 
  relocate(wa_max_n:wa_max_tot, .before = bws_n) |>
  relocate(wp_max_n:wp_max_agr, .before = cep_n) |> 
  mutate(across(where(is.numeric), ~ if_else(between(.x, 1, 5), .x, NA)))
```

```{r}
st_drop_geometry(son_water_adm0) |> 
  skim()
```

### 5.2. Sub-National level

```{r}
bws_adm1 <- aqueduct_adm1 |> 
  filter(indicator_name == "bws") |> 
  mutate(
    weight = if_else(weight == "One", "n", tolower(weight)),
    score = na_if(score, -9999),
    # rescale values 0-5 to 1-5 applying (new_max - new_min) * (x - min_old) / (max_old - min_old) + min_new
    score = (5-1) * ((score-0) / (5-0)) + 1 
  ) |> 
  left_join(GADM_trans |> select(GID_1, ADM1CD_c), by = c("gid_1" = "GID_1")) |> 
  filter(!is.na(ADM1CD_c)) |> ## Some cases where there is no match (disputed)
  select(ADM1CD_c, GID_1 = gid_1, weight, score) |> 
  pivot_wider(names_from = weight, values_from = score, names_prefix = "bws_") |> 
  relocate(ADM1CD_c, GID_1, bws_n, bws_dom, bws_ind, bws_irr, bws_liv, bws_tot)
```

```{r}
son_water_adm1_extract <- adm1 |>
  # baseline water stress (bws)
  left_join(bws_adm1, by = "ADM1CD_c") |> 
  # water depletion (wdp)
  left_join(exact_extract(list_wa$wdp_n, adm1, fun = "mean", append_cols = "ADM1CD_c", progress = TRUE) |> rename(wdp_n = mean), by = "ADM1CD_c") |>
  left_join(aggregate_to_adm(list_wa$wdp_n, wdp_weights$dom, wdp_dom, adm1, "ADM1CD_c"), by = "ADM1CD_c") |>
  left_join(aggregate_to_adm(list_wa$wdp_n, wdp_weights$ind, wdp_ind, adm1, "ADM1CD_c"), by = "ADM1CD_c") |>
  left_join(aggregate_to_adm(list_wa$wdp_n, wdp_weights$irr, wdp_irr, adm1, "ADM1CD_c"), by = "ADM1CD_c") |>
  left_join(aggregate_to_adm(list_wa$wdp_n, wdp_weights$liv, wdp_liv, adm1, "ADM1CD_c"), by = "ADM1CD_c") |>
  left_join(aggregate_to_adm(list_wa$wdp_n, wdp_weights$tot, wdp_tot, adm1, "ADM1CD_c"), by = "ADM1CD_c") |>
  # freshwater quantity target (fqt)
  left_join(exact_extract(list_wa$fqt_n, adm1, fun = "mean", append_cols = "ADM1CD_c", progress = TRUE) |> rename(fqt_n = mean), by = "ADM1CD_c") |>
  left_join(aggregate_to_adm(list_wa$fqt_n, fqt_weights$agr, fqt_agr, adm1, "ADM1CD_c"), by = "ADM1CD_c") |>
  left_join(aggregate_to_adm(list_wa$fqt_n, fqt_weights$dom, fqt_dom, adm1, "ADM1CD_c"), by = "ADM1CD_c") |>
  left_join(aggregate_to_adm(list_wa$fqt_n, fqt_weights$ind, fqt_ind, adm1, "ADM1CD_c"), by = "ADM1CD_c") |>
  left_join(aggregate_to_adm(list_wa$fqt_n, fqt_weights$tot, fqt_tot, adm1, "ADM1CD_c"), by = "ADM1CD_c") |> 
  # coastal eutrophication potential (cep)
  left_join(exact_extract(list_wp$cep_n, adm1, fun = "mean", append_cols = "ADM1CD_c", progress = TRUE) |> rename(cep_n = mean), by = "ADM1CD_c") |>
  left_join(aggregate_to_adm(list_wp$cep_n, wp_weights, cep_agr, adm1, "ADM1CD_c"), by = "ADM1CD_c") |>
  # nitrate-nitrite concentration (nox)
  left_join(exact_extract(list_wp$nox_n, adm1, fun = "mean", append_cols = "ADM1CD_c", progress = TRUE) |> rename(nox_n = mean), by = "ADM1CD_c") |>
  left_join(aggregate_to_adm(list_wp$nox_n, wp_weights, nox_agr, adm1, "ADM1CD_c"), by = "ADM1CD_c") |>
  # periphyton growth potential (pgp)
  left_join(exact_extract(list_wp$pgp_n, adm1, fun = "mean", append_cols = "ADM1CD_c", progress = TRUE) |> rename(pgp_n = mean), by = "ADM1CD_c") |>
  left_join(aggregate_to_adm(list_wp$pgp_n, wp_weights, pgp_agr, adm1, "ADM1CD_c"), by = "ADM1CD_c")

son_water_adm1 <- son_water_adm1_extract |> 
  mutate(across(where(is.numeric), ~ round(.x, 1))) |> 
  rowwise() |> 
  mutate(
    wa_max_n = max(c(bws_n, wdp_n, fqt_n), na.rm = TRUE),
    wa_max_dom = max(c(bws_dom, wdp_dom, fqt_dom), na.rm = TRUE),
    wa_max_ind = max(c(bws_ind, wdp_ind, fqt_ind), na.rm = TRUE),
    wa_max_irr = max(c(bws_irr, wdp_irr, fqt_agr), na.rm = TRUE),
    wa_max_liv = max(c(bws_liv, wdp_liv, fqt_n), na.rm = TRUE),
    wa_max_tot = max(c(bws_tot, wdp_tot, fqt_tot), na.rm = TRUE),
    wp_max_n = max(c(cep_n, nox_n, pgp_n), na.rm = TRUE),
    wp_max_agr = max(c(cep_agr, nox_agr, pgp_agr), na.rm = TRUE)
  ) |> 
  ungroup() |> 
  relocate(wa_max_n:wa_max_tot, .before = bws_n) |>
  relocate(wp_max_n:wp_max_agr, .before = cep_n) |> 
  mutate(across(where(is.numeric), ~ if_else(between(.x, 1, 5), .x, NA)))
```

```{r}
st_drop_geometry(son_water_adm1) |> 
  skim()
```

## 6. Export

```{r}
key_descriptions <- tribble(
  ~Key,    ~Description,     
  "wa_max",    "Maximum value between water availability layers: bws, wdp, fqt", 
  "wp_max",  "Maximum value between water pollution layers: cep, nox, pgp", 
  "bws","Baseline Water Stress",     
  "wdp","Water depletion",      
  "fqt","Freshwater Quantity Target",
  "cep","Coastal Eutrophication Potential",
  "nox","Nitrate-Nitrite Concentration",
  "pgp","Periphyton Growth Potential",
  "","",
  "","",
  "Suffix","Description",
  "_n","Aggregation at administrative level, without weighting",
  "_dom","Aggregation at administrative level weighted by domestic water demand*",
  "_ind","Aggregation at administrative level weighted by industrial water demand*",
  "_irr","Aggregation at administrative level weighted by irrigation water demand*",
  "_liv","Aggregation at administrative level weighted by livestock water demand*",
  "_tot","Aggregation at administrative level weighted by total water demand*",
  "_agr","Aggregation at administrative level weighted by cropland extent**",
  "","",
  "", "* Data for 'water demand' differs depending on the indicator aggregated",
  "bws","Water withdrawal from PCR-GLOBWB 2",
  "wdp","Water consumption from WaterGAP 3",
  "fqt","Freshwater Quantity Target from Water Footprint Network",
  "", "** For the water pollution indicators we used cropland extend from Potapov et al. 2021 (https://www.nature.com/articles/s43016-021-00429-z)"
)
```

```{r}
write.xlsx(list(
  "Help" = key_descriptions,
  "National level" = st_drop_geometry(son_water_adm0) |> 
    mutate(notes = if_else(LOCATION_ID %in% c("RUS", "CAN", "CHN", "USA", "BRA", "AUS", "IND", "ARG", "KAZ", "DZA"),
                           "Not recommended to be used. Average values for such large country are not representative. Use values from sub-national level instead.", NA)), 
  "Sub-national level" = st_drop_geometry(son_water_adm1)
), file = here("v2.1", "outputs", "adm_level", "sbtn_son_water_v2.1_adm_level.xlsx"), rowNames = FALSE)
```

*National level*

```{r}
write_rds(son_water_adm0, here(params$version, "outputs", "adm_level", "sbtn_son_water_v2.1_adm0.rds"))
```

*Sub-national level*

```{r}
write_rds(son_water_adm1, here(params$version, "outputs", "adm_level", "sbtn_son_water_v2.1_adm1.rds"))
```

## 7. View outputs

```{r}
son_water_adm0 <- read_rds(here("v2.1", "outputs", "adm_level", "sbtn_son_water_v2.1_adm0.rds"))
son_water_adm1 <- read_rds(here("v2.1", "outputs", "adm_level", "sbtn_son_water_v2.1_adm1.rds"))
```

```{r}
mapview_custom <- function(data, layer_name, var){
  mapview(
    data,
    layer.name = layer_name,
    zcol = var,
    col.regions = colorRampPalette(c("#FFFF99", "#FFE600", "#FF9900", "#FF1900", "#990000")),
    at = c(1, 1.8, 2.6, 3.4, 4.2, 5),
    lwd = 0.1
  )
}
```

*National level*

```{r}
son_water_adm0_simplified <- son_water_adm0 |> 
  st_transform("EPSG:8857") |> 
  st_simplify(dTolerance = 1000) |> 
  st_transform("EPSG:4326")
```

```{r}
mapview_custom(son_water_adm0_simplified, "Water Availability (Max)", "wa_max_tot")
```

```{r}
mapview_custom(son_water_adm0_simplified, "Water Pollution (Max)", "wp_max_agr")
```

*Sub-national level*

```{r}
son_water_adm1_simplified <- son_water_adm1 |> 
  st_transform("EPSG:8857") |> 
  st_simplify(dTolerance = 1000) |> 
  st_transform("EPSG:4326")
```

```{r}
mapview_custom(son_water_adm1_simplified, "Water Availability (Max)", "wa_max_tot")
```

```{r}
mapview_custom(son_water_adm1_simplified, "Water Pollution (Max)", "wp_max_agr")
```
